---
title: "How buffer size and igotu fix length influence behavior analysis"
author: "Amy Van Scoyoc"
date: "5/20/2020"
output: html_document
---

#Load Packages
```{r}
#load packages
library(tidyverse)
library(plyr)
library(readr)
library(amt)
library(survival)
```

#Import

First, I import the raw igotu data for all animals, combine the .csvs, and add a column for the animal ID based on the file name. I then format the timestamps and nest the data by ID so that we can create a track for each animal using the `amt` package.  
```{r}
#set working directory
setwd("/Users/Amy/Berkeley Classes/Brashares Lab/projects/igotu/raw_data")
#set options to prevent truncated lat/long
options(digits=15)
#get file names
files <- list.files()
#read in files
data <- purrr::map_df(files, function(x) {
                      	data <- read_csv(x) %>% select(1:4)
                      	cbind(ID = substr(x,1,2), data)
}) %>% #create timestamp
  mutate(timestamp = as.POSIXct(paste(Date, Time), format = "%Y-%m-%d %H:%M:%S", tz = "America/Los_Angeles")) %>% 
  nest(-ID) #nest the data by animal ID
```

Next, we create a track for each individual animal using the `amt` package and transform the coordinates to UTMs so that it is easier to work with.
```{r}
#make "track" for each ID using amt package
igotu <- data %>% 
  mutate(trk = map(data, function(d){
    mk_track(d, Longitude, Latitude, timestamp, crs = CRS("+init=epsg:4326")) %>% 
      # Transform latlongs into UTMs
      transform_coords(CRS("+proj=utm +zone=10 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"))
  }))
```

#Resampling

In order to measure the impact of fix rate on fence crossing analyses, we have to resample the animal tracks to various intervals. 

First, I check the sampling rate of the raw data, just to see what we are working with. 
```{r}
#check sampling rate
igotu %>%
  mutate(sr = lapply(trk, summarize_sampling_rate)) %>% 
  dplyr::select(ID, sr) %>% unnest
```

Next, I write a function to classify animal tracks into bursts, eliminating bursts with fewer than 2 points. I run this function with various intervals to create data frames with bursts of different fix rates.  *Note that this function uses `seconds()` as the rate, so multiply by 60 to convert to minutes and again by 60 to convert to hours.  
```{r}
#resample function for multiple intervals (in seconds)
resample <- function(df, rate){
            steps <- df %>% 
                mutate(step = map(df[[3]],function(x){
                              x %>% 
                              # Eliminate steps longer than 3 hours
                              track_resample(rate = seconds(rate), tolerance = seconds(1)) %>%
                              # Eliminate bursts with fewer than 2 points
                              filter_min_n_burst(min_n = 2)}))
            return(steps)
}
#apply resample function
ten_sec <- resample(igotu,10)
five_min <- resample(igotu,5*60)
thirty_min <- resample(igotu,30*60)
one_hr <- resample(igotu,1*60*60)
three_hr <- resample(igotu,3*60*60)
three_hr[[4]]
#combine dfs into single list for interation later
frames <- list(ten_sec,five_min,thirty_min,one_hr,three_hr)
```

#Fences
Here, I load in fence shapefiles of HREC using the `sf` package. I specify fence buffer sizes I want to test and transform the crs into UTMs to match the animal data. 
```{r}
library(sf)

#read in fence shapefile
fences <- st_read("/Users/Amy/Desktop/igotu/fences/Fences_103015.shp")
#set crs 
crs = "+proj=utm +zone=10 +datum=WGS84 +units=m +no_defs"
#adjust buffer distances 
fence_buff_500m <- st_buffer(fences, 500) %>% st_transform(crs = st_crs(crs))
fence_buff_100m <- st_buffer(fences, 100) %>% st_transform(crs = st_crs(crs))
fence_buff_50m <- st_buffer(fences, 50) %>% st_transform(crs = st_crs(crs))
fence_buff_10m <- st_buffer(fences, 10) %>% st_transform(crs = st_crs(crs))
fence_buff_1m <- st_buffer(fences, 1) %>% st_transform(crs = st_crs(crs))
fence_buff_.5m <- st_buffer(fences, 0.5) %>% st_transform(crs = st_crs(crs))
```

#Fence Encounters
To measure whether an animal crosses a fence, I create a function that counts the number of bursts that intersect a fence buffer. 
```{r}
#create encounter function
encounters <- function(x, crs, layer){
  temp <- map(x$step, function(y){
                                y %>% 
                                  st_as_sf(coords = c("x_", "y_")) %>% 
                                  group_by(burst_) %>% 
                                  summarize(do_union = FALSE) %>% 
                                  st_cast("LINESTRING") %>% 
                                  st_set_crs(crs) %>% 
                                  st_intersects(layer) %>% 
                                  map(length) %>% 
                                  subset(. != 0) %>% 
                                  length()}) %>% 
          unlist() %>% 
          `names<-`(as.character(quote(x))) %>% 
          as_tibble()
  return(temp)
}
```

I apply this function to all animals frames, with each combination of fix rate and buffer size. 
```{r}
#set new inits
#frames <- list(tensec_dat,fivemin_dat,thirtymin_dat,onehour_dat,threehour_dat)
fixes <- c("10-sec","5-min", "30-min", "1-hr", "3-hr") %>% rep(each=4)
animal_IDs <- igotu[[1]]
#run encounter function
n_crossings <- list("500-m" = map_dfc(frames, encounters, crs, fence_buff_500m),
                    "100-m" = map_dfc(frames, encounters, crs, fence_buff_100m),
                    "50-m" = map_dfc(frames, encounters, crs, fence_buff_50m), 
                    "10-m" = map_dfc(frames, encounters, crs, fence_buff_10m), 
                    "1-m" = map_dfc(frames, encounters, crs, fence_buff_1m),
                    ".5-m" = map_dfc(frames, encounters, crs, fence_buff_.5m)) %>% 
               bind_rows(.id="buffers") %>% 
               cbind(ID = animal_IDs,.) %>% 
               `names<-`(colnames)
n_crossings


cross <- list("500-m" = map_dfr(frames, encounters, crs, fence_buff_500m),
              "100-m" = map_dfr(frames, encounters, crs, fence_buff_100m),
              "50-m" = map_dfr(frames, encounters, crs, fence_buff_50m), 
              "10-m" = map_dfr(frames, encounters, crs, fence_buff_10m), 
              "1-m" = map_dfr(frames, encounters, crs, fence_buff_1m),
              ".5-m" = map_dfr(frames, encounters, crs, fence_buff_.5m)) %>% 
              bind_rows(.id="buffers") %>% 
              cbind(ID = animal_IDs,fixes = fixes,.)

#length(which(sapply(., length)!=0))
#length(which(sapply(, length)!=0))/length(is)*100
#plot(st_geometry(fence_buff_100m))
#plot(st_geometry(is),col="red", add=T)
```

To ensure that we compare the proportion of fence crossings at each fix rate (i.e. higher fix rates have higher numbers of bursts, so the numbers are somewhat misleading to compare), I create a function to total the number of bursts for each animal ID at each fix rate.  I bind this to our data frame and use it to calculate the proportion of crossings. 
```{r}
#calculate total # of steps (i.e. bursts) per track
totals <- function(x){
      t <- map(x$step, function(y){
                  y %>% 
                  group_by(burst_) %>% 
                  tally() %>%
                  nrow()}) %>% 
      unlist() %>% 
      `names<-`(as.character(quote(x))) %>% 
      as_tibble()
  return(t)
}

t <- list("totals" = map_dfr(frames, totals)) %>% 
  bind_rows() %>%
  cbind(id=paste0(animal_IDs,fixes),ID=animal_IDs, fixes=fixes,.) %>% 
  select(id=id,total_bursts=4)
     
n_cross <- cross %>% 
  mutate(id=paste0(ID,fixes)) %>% 
  left_join(.,t,by="id") %>% 
  mutate(percent_cross = value/total_bursts*100) %>% 
  `names<-`(c("ID", "fix", "fence_buffer","cross_bursts","id","total_bursts","percent_cross"))

```

#Plotting
Here are the results: 

```{r}
#plot effect of fence buffer and fix-length on # of steps encountering fence
ggplot(n_cross, aes(x=fix, y=cross_bursts)) +
  scale_x_discrete(limits = c("10-sec","5-min", "30-min", "1-hr", "3-hr")) + 
  geom_line(aes(group=interaction(ID, fence_buffer), color=fence_buffer)) +
  ggtitle("Effect of fix length and buffer size on fence encounter bursts") +
  ylab("Fence encounters (bursts)") +
  xlab("Fix length") +
  theme_classic()
```


```{r}
#plot median of fence buffer and fix-length on # of steps encountering fence
median_n_bursts <- n_cross %>%
                group_by(fence_buffer,fix) %>%
                summarise(median = median(cross_bursts))

ggplot(median_n_bursts, aes(x=fix, y=median)) +
  scale_x_discrete(limits = c("10-sec","5-min", "30-min", "1-hr", "3-hr")) + 
  geom_line(aes(group=fence_buffer,color=fence_buffer)) +
  ggtitle("Median effect of fix length and buffer size on fence encounter bursts (N = 4 deer)") +
  ylab("Fence encounters (bursts)") +
  xlab("Fix length") +
  theme_classic()
```

```{r}
#this is misleading though because I used #of bursts
#we need to look at the %encounters (since #bursts decreases with increasing fix rate)
#to do this, I add a column to with the total #bursts, then I calculate %bursts
ggplot(n_cross, aes(x=fix, y=percent_cross)) +
  scale_x_discrete(limits = c("10-sec","5-min", "30-min", "1-hr", "3-hr")) + 
  geom_line(aes(group=interaction(ID, fence_buffer), color=fence_buffer)) +
  ggtitle("Effect of fix length on % fence encounters (bursts)") +
  ylab("Fence encounters (bursts)") +
  xlab("Fix length") +
  theme_classic()
```


```{r}
#plot median of fence buffer and fix-length on # of steps encountering fence
median_perc_bursts <- n_cross %>%
                group_by(fence_buffer,fix) %>%
                summarise(median = median(percent_cross))

ggplot(median_perc_bursts, aes(x=fix, y=median)) +
  scale_x_discrete(limits = c("10-sec","5-min", "30-min", "1-hr", "3-hr")) + 
  geom_line(aes(group=fence_buffer,color=fence_buffer)) +
  ggtitle("Median effect of fence buffer size on percent fence encounters (N = 4 deer)") +
  ylab("Fence encounters (bursts)") +
  xlab("Fix length") +
  theme_classic()
```


```{r}
ggplot(n_cross, aes(x=fence_buffer, y=percent_cross)) +
  scale_x_discrete(limits = c(".5-m","1-m", "10-m", "50-m", "100-m","500-m")) + 
  geom_line(aes(group=interaction(ID, fix), color=fix)) +
  #facet_grid(ID~.)+
  ggtitle("Effect of fix rate on % fence encounters (bursts)") +
  ylab("Fence encounters (bursts)") +
  xlab("Fence buffer (meters)") +
  theme_classic()
```


```{r}
#plot median of fence buffer and fix-length on # of steps encountering fence
ggplot(median_perc_bursts, aes(x=fence_buffer, y=median)) +
  scale_x_discrete(limits = c(".5-m","1-m", "10-m", "50-m", "100-m","500-m")) + 
  geom_line(aes(group=fix, color=fix)) +
  #facet_grid(ID~.)+
  ggtitle("Effect of fixe rate on % fence encounters (bursts)") +
  ylab("Fence encounters (bursts)") +
  xlab("Fence buffer (meters)") +
  theme_classic()
```


This looks great, now I just need to filter the igotu data so that the fixes are only on the deer after release (some pre-data from HREC and stuff is there which influences encounter rates in the bursts).  See igotu_exploratory.rmd



